\textit{Red Oak}\\ Proactive Controller Generation for Algorithmic Policy

### 2015-05-12 draft.tex
..

\documentclass[preprint]{sigplanconf}
\usepackage[latin1]{inputenc}
\usepackage{times}
\usepackage{latexsym}
\usepackage{graphicx}
\usepackage{enumitem}
\usepackage{listings}
\usepackage{bussproofs}
\usepackage{comment}
\usepackage[numbers]{natbib}
\usepackage{graphicx} % Graphics support
\usepackage{caption} % Improved captions
\usepackage{subfig} % Inclusion of subfigures

@@

\section*{Algorithic Policies Revisited}
% Explains what algorithmic policies are, how maple was handling them and the difference with our model 
Algorithmic policies were introduced by Maple as a model for SDN controllers. An algorithmic policy is a programmer-defined function from a packet to a forwarding decision (i.e., where the packet(s) should end up). Conceptually this model offers the abstraction that the network forwards packets as if this function were executed on each packet when entering the network. To handle future traffic directly in the data plane, Maple keeps a tree data structure that branches on decisions observed at run time (e.g., the packet had source IP 10.0.0.1). We extend the idea of decision tree to work at a higher level of abstraction involving decisions based on symbolic constraints, rather than concrete, packets. Let us take a look at an example.

@@

   We provide several types of forwarding decisions to the programmer. The programmer can deliver a packet, according to the packet's desination, drop a packet, and forwarding a packet to a static location. Our API also allows for dynamic forwarding based on component of relation, for example to the last seen location of a host according to a MAC-learning controller (not shown in Fig~\ref{fig:range_api}). Finally, the programmer can combine several forwarding decisions into one by using a mulitcast forwarding decision.

   \subsection*{Combining Algorithmic Policies}
   % cross, ``holes''
   As algorithmic policies under our model are functions of the host language, we inherit the latter's function space ability, allowing, amongst other things, for compositon of algorithmic policies and higher-order algorithmic policies.

@@

\begin{figure}[ht]
\begin{lstlisting}[mathescape]
let compile (f: pkt $\rightarrow$ decision) : policy =
      stack $\gets$ empty.push(new packet)
      root $\gets$ new unexplored node
      while (stack is not empty)
          loc $\gets$ root
          pkt $\gets$ stack.pop
          fd $\gets$ f(pkt) 
          loc $\gets$ Leaf fd
      return root    
\end{lstlisting}
\caption{Pseudocode for the compile function}
\label{fig:compile-pseudo}
\end{figure}

To do so, we repeatedly run the algorithmic policy \lstinline|f| on symbol packets generated to explore every reachable path.
We begin by calling \lstinline|f| on a symbolic packet covering the full range of value for each of its fields.

@@

Similarly, when the algorithmic policy calls \lstinline|in_relation| to learn if fields \lstinline|flds|'s value are in a relation \lstinline|rel|, we
\begin{enumerate}
  \item Create a new \lstinline|in_rel| branching node in the decision tree wit hsubtrees corresponding to \lstinline|true| and \lstinline|false|
  \item Tag the current packet with the symbolic constraint of being in the relation and return true
  \item Push onto a stack consisting of remaining packets to process, a new symbolic that is the same as the current packet, but tagged as not being part of the relation \lstinline|rel|.
\end{enumerate}
When \lstinline|f| calls \lstinline|add| or \lstinline|remove|, we record in the decision the current symbolic packet, the relation and the fields being added or removed from the relation before continuing with the execution of \lstinline|f|.

@@

When \lstinline|compile| terminates, a complete decision tree has been traced for the input algoritmic policy. The next step is generate symbolic rules, mapping a symbolic packet to a forwarding decision.
To do so, we do a depth-first, left to right traversal of the tree. On reaching an \lstinline|Add| or a \lstinline|Remove| node, we create a new rule sending the relevant packets to the controller, pruning the rest of the sub-tree. On reaching a leaf, we create a new rule forwarding relevant packets according to the forwarding decision held in the leaf node.

We then generate concrete rules from a collection of relations and a list of symbolic rules. Concrete rules differ from symbolic rules in that they do not depend on relations. Each symbolic rule expands into a list of concrete rules. For each tuple in the relation on which the symbolic rules depend, we create a new rule, refining the ranges of the relevant fields of the symbolic packet with the values of the tuple. We repeat this operation on the created list of symbolic rules until all dependant relations are taken care of, at which point the rules are considered concrete. Due to the prioritization of the resulting rules, we do not need to modify symbolic rules marked as not in a certain relation, as all the packets in the relation will have been matched by the positive rule before reaching this one.

@@

  Assume that the trusted connection relation ``conn'' contains the pairs 
  \lstinline| {(150,10), (200, 9), (8, 150)}|, meaning external Src 150 can talk to internal Dst 10 and so on. We concretize the above symbolic rules by expanding the first rule to:
\begin{lstlisting}[mathescape]
Src $\in$ ([101, 1000] $\cap$ 150), Dst = 10 $\mapsto$ Deliver
Src $\in$ ([101, 1000] $\cap$ 200), Dst = 9 $\mapsto$ Deliver
Src $\in$ ([101, 1000] $\cap$ 8), Dst = 150 $\mapsto$ Deliver
 \end{lstlisting}
To satisfy both symbolic constraints, we take the intersection of the constraints, which corresponds to intersection of the constraints on the ranges of each field. These rules reduce to:
\begin{lstlisting}[mathescape]
Src = 150, Dst = 10 $\mapsto$ Deliver
Src = 200, Dst = 9 $\mapsto$ Deliver
 \end{lstlisting}
Note the omission of the third concrete rule, resulting from \lstinline|Src| having an empty range. We can now concretize the second symbolic rule without doing any work, since checking for relation membership is already done by all the higher priority rules. The final concrete forwarding rules are now:
\begin{lstlisting}[mathescape]
Src = 150, Dst = 10 $\mapsto$ Deliver
Src = 200, Dst = 9 $\mapsto$ Deliver
Src $\in$ [101, 1000] $\mapsto$ Drop
\end{lstlisting}

@@

\begin{figure}[ht]
\begin{lstlisting}[mathescape]
let run (pol: policy) (inputs: examples) : unit =
      sr $\gets$ build_symbolic_rules(pol)
      rel $\gets$ empty
      cr $\gets$ build_concrete_rules(sr, rel)
      for each packets pkt in input,
      match evaluate_rules(cr, pkt) with
       | Some fd $\to$ forward(pkt, fd)
       | None $\to$
           (fd, rel) $\gets$ evaluate_policy(pkt)
           cr $\gets$ build_concrete_rules(sr, rel)
           forward(pkt, fd)     
\end{lstlisting}

@@

We implemented in \textit{Red Oak} a local transformation adding an \lstinline|in_rel| node before \lstinline|add| or \lstinline|remove| nodes. In the case of \lstinline|add|, the \lstinline|add| node is omited in the true branch of \lstinline|in_rel|, such that packets whose fields are already in the relation will not be seen by the controller due to this \lstinline|add| node. Similarly, we omit the \lstinline|remove| node in the false branch of \lstinline|in_rel|. In the firewall example (see Fig.~\ref{fig:firewallcode}), this would result the following decision tree:

>>>>

\documentclass[nocopyrightspace]{sigplanconf}
\usepackage[latin1]{inputenc}
\usepackage{times}
\usepackage{latexsym}
\usepackage{graphicx}
\usepackage{enumitem}
\usepackage{listings}
\usepackage{bussproofs}
\usepackage{comment}
\usepackage[numbers]{natbib}
\usepackage{graphicx} % Graphics support
\usepackage{caption} % Improved captions
\usepackage{subfig} % Inclusion of subfigures

@@

\section*{Algorithmic Policies Revisited}
% Explains what algorithmic policies are, how maple was handling them and the difference with our model 
Algorithmic policies were introduced by Maple as a model for SDN controllers. An algorithmic policy is a programmer-defined function from a packet to a forwarding decision (i.e., where the packet(s) should end up). Conceptually this model offers the abstraction that the network forwards packets as if this function were executed on each packet when entering the network. To handle future traffic directly in the data plane, Maple keeps a tree data structure that branches on decisions observed at run time (e.g., the packet had source IP 10.0.0.1). We extend the idea of decision tree to work at a higher level of abstraction involving decisions based on symbolic constraints, rather than concrete, packets. Let us take a look at an example.

@@

   We provide several types of forwarding decisions to the programmer. The programmer can deliver a packet, according to the packet's destination, drop a packet, and forwarding a packet to a static location. Our API also allows for dynamic forwarding based on component of relation, for example to the last seen location of a host according to a MAC-learning controller (not shown in Fig~\ref{fig:range_api}). Finally, the programmer can combine several forwarding decisions into one by using a mulitcast forwarding decision.

   \subsection*{Combining Algorithmic Policies}
   % cross, ``holes''
   As algorithmic policies under our model are functions of the host language, we inherit the latter's function space ability, allowing, among other things, for composition of algorithmic policies and higher-order algorithmic policies.

@@

\begin{figure}[ht]
\begin{lstlisting}[mathescape]
let compile (f: pkt $\rightarrow$ decision) : policy =
      stack $\gets$ empty.push(new sym packet)
      root $\gets$ new unexplored node
      while (stack is not empty)
          loc $\gets$ root
          pkt $\gets$ stack.pop
          fd $\gets$ f(pkt) 
          loc $\gets$ Leaf fd
      return root    
\end{lstlisting}
\caption{Pseudocode for the compile function}
\label{fig:compile-pseudo}
\end{figure}

To do so, we repeatedly run the algorithmic policy \lstinline|f| on symbol packets generated to explore every reachable path.
We begin by calling \lstinline|f| on a symbolic packet covering the full range of values for each of its fields.

@@

Similarly, when the algorithmic policy calls \lstinline|in_relation| to learn if fields \lstinline|flds|'s value are in a relation \lstinline|rel|, we
\begin{enumerate}
  \item Create a new \lstinline|in_rel| branching node in the decision tree with subtrees corresponding to \lstinline|true| and \lstinline|false|
  \item Tag the current packet with the symbolic constraint of being in the relation and return true
  \item Push onto a stack consisting of remaining packets to process, a new symbolic that is the same as the current packet, but tagged as not being part of the relation \lstinline|rel|.
\end{enumerate}
When \lstinline|f| calls \lstinline|add| or \lstinline|remove|, we record in the decision the current symbolic packet, the relation and the fields being added or removed from the relation before continuing with the execution of \lstinline|f|.

@@

When \lstinline|compile| terminates, a complete decision tree has been traced for the input algorithmic policy. The next step is generate symbolic rules, mapping a symbolic packet to a forwarding decision.
To do so, we do a depth-first, left to right traversal of the tree. On reaching an \lstinline|Add| or a \lstinline|Remove| node, we create a new rule sending the relevant packets to the controller, pruning the rest of the sub-tree. On reaching a leaf, we create a new rule forwarding relevant packets according to the forwarding decision held in the leaf node.

We then generate concrete rules from a collection of relations and a list of symbolic rules. Concrete rules differ from symbolic rules in that they do not depend on relations. Each symbolic rule expands into a list of concrete rules. For each tuple in the relation on which the symbolic rules depend, we create a new rule, refining the ranges of the relevant fields of the symbolic packet with the values of the tuple. We repeat this operation on the created list of symbolic rules until all dependent relations are taken care of, at which point the rules are considered concrete. Due to the prioritization of the resulting rules, we do not need to modify symbolic rules marked as not in a certain relation, as all the packets in the relation will have been matched by the positive rule before reaching this one.

@@

  Assume that the trusted connection relation ``conn'' contains the pairs 
  \lstinline| {(150,10), (200, 9), (8, 150)}|, meaning external Src 150 can talk to internal Dst 10 and so on. We make the above symbolic rules concrete by expanding the first rule to:
\begin{lstlisting}[mathescape]
Src $\in$ ([101, 1000] $\cap$ 150), Dst = 10 $\mapsto$ Deliver
Src $\in$ ([101, 1000] $\cap$ 200), Dst = 9 $\mapsto$ Deliver
Src $\in$ ([101, 1000] $\cap$ 8), Dst = 150 $\mapsto$ Deliver
 \end{lstlisting}
To satisfy both symbolic constraints, we take the intersection of the constraints, which corresponds to intersection of the constraints on the ranges of each field. These rules reduce to:
\begin{lstlisting}[mathescape]
Src = 150, Dst = 10 $\mapsto$ Deliver
Src = 200, Dst = 9 $\mapsto$ Deliver
 \end{lstlisting}
Note the omission of the third concrete rule, resulting from \lstinline|Src| having an empty range. We can now make the second symbolic rule concrete without doing any work, since checking for relation membership is already done by all the higher priority rules. The final concrete forwarding rules are now:
\begin{lstlisting}[mathescape]
Src = 150, Dst = 10 $\mapsto$ Deliver
Src = 200, Dst = 9 $\mapsto$ Deliver
Src $\in$ [101, 1000] $\mapsto$ Drop
\end{lstlisting}

@@

\begin{figure}[ht]
\begin{lstlisting}[mathescape]
let run (pol: policy) (inputs: examples) : unit =
      sr $\gets$ build_symbolic_rules(pol)
      rel $\gets$ empty
      cr $\gets$ build_concrete_rules(sr, rel)
      for each packets pkt in input,
      match evaluate_rules(cr, pkt) with
       | Some fd $\to$ forward(pkt, fd)
       | None $\to$
         (fd, rel) $\gets$ evaluate_policy(pol, pkt)
         cr $\gets$ build_concrete_rules(sr, rel)
         forward(pkt, fd)     
\end{lstlisting}

@@

We implemented in \textit{Red Oak} a local transformation adding an \lstinline|in_rel| node before \lstinline|add| or \lstinline|remove| nodes. In the case of \lstinline|add|, the \lstinline|add| node is omitted in the true branch of \lstinline|in_rel|, such that packets whose fields are already in the relation will not be seen by the controller due to this \lstinline|add| node. Similarly, we omit the \lstinline|remove| node in the false branch of \lstinline|in_rel|. In the firewall example (see Fig.~\ref{fig:firewallcode}), this would result the following decision tree:
